---
title: "Untitled"
author: "Sean Ng"
date: "11/13/2021"
output: 
  html_document:
    code_download: true
    theme: readable
---

```{=html}
<script src="hideOutput.js"></script>
```
```{css, echo=FALSE}
.showopt {
  background-color: #004c93;
  color: #FFFFFF; 
  width: 100px;
  height: 20px;
  text-align: center;
  vertical-align: middle !important;
  float: right;
  font-family: sans-serif;
  border-radius: 8px;
}

.showopt:hover {
    background-color: #dfe4f2;
    color: #004c93;
}

pre.plot {
  background-color: white !important;
}
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=9, message = FALSE, warning=FALSE)
library(tidyverse)
library(readxl)
library(lubridate)
library(janitor)
library(stringi)
library(stringr)

library(pander)
library(knitr)
library(kableExtra)
library(ggmap)
library(sf)
library(plotly)
library(scales)
library(ggforce)
library(ggpubr)
library(forcats)
library(patchwork)
library(rattle)
library(rpart)
library(rpart.plot)
library(rpart.utils)
library(partykit)
library(corrplot)
library(dendextend)
library(factoextra)

# disabling scientific notation
options(scipen = 100)

# pander tables all in one row
panderOptions('table.split.table', Inf)

# pander thousands separator
panderOptions("big.mark", ",")
```

::: {.fold .s .o}
```{r reading-and-cleaning}

# function to remove accents 
rm_accent <- function(colns){
  colns <- stri_trans_general(colns, "Latin-ASCII")
}

# reading and cleaning -- you really should break it into parts
ven1 <- read_csv("consolidation 191209 1635.csv") %>% 
  clean_names() %>% 
  # removing unused columns
  select(-c(codigodeestablecimientoocentro, loc_id, hrp_sitre_p_indicator, 
            tipoderespuesta, comentarios, coordeadas_gps_x, coordeadas_gps_y,
            fechade_inicio, fecha_previstade_finalizacion)) %>% 
  # renaming unwieldy columns 
  rename(ubicacion          = comunidadonombredelestablecimiento_centro, 
         sector             = sector_areade_responsabiliad,
         beneficiarios_meta = beneficiarios_meta_numerodepersonas,
         estatus            = estatusdeprogramacion) %>% 
      # mutating the date to the right format
  mutate(month = as.factor(recode(month,
                        `4` = "30/04/2019",
                        `5` = "31/05/2019",
                        `6` = "30/06/2019",
                        `7` = "31/07/2019",
                        `8` = "31/08/2019",
                        `9` = "30/09/2019",
                        `10` = "31/10/2019"))) %>% 
  mutate(month = as.Date(month %>% strptime(., format = "%d/%m/%Y"))) %>% 
  mutate(org_lider = coalesce(org_lider, org_implementadora)) %>% 
  # correcting sector names
  mutate(sector = str_replace_all(sector, c(
    "Agua_saneamiento_higiene"            = "WASH",
    "educacion"                           = "Educacion",
    "Nutricion"                           = "Nutricion",
    "protección_Niños_Niñas_Adolescentes" = "Proteccion_NNA",
    "Protección_Niños_Niñas_Adolescentes" = "Proteccion_NNA",
    "Protección_Violencia_Género"         = "Proteccion_GBV"))) %>% 
  # renaming beneficiary disaggregation columns 
  rename(f_0_18 = f_18,
         m_0_18 = m_18,
         f_18plus = f_18_2,
         m_18plus = m_18_2) %>% 
  mutate(estado    = rm_accent(str_to_upper(estado)), 
         municipio = rm_accent(str_to_upper(municipio)),
         parroquia = rm_accent(str_to_upper(parroquia)),
         ubicacion = rm_accent(str_to_upper(ubicacion)),
         actividad = rm_accent(str_to_upper(actividad)),
         categoria = rm_accent(str_to_upper(categoriadeactividad))) %>% 
  # recoding the estatus column 
  mutate(estatus = str_replace_all(estatus, 
                  c("En ejecucion" = "ejecucion", 
                    "en ejecución" = "ejecucion", 
                    "en Ejecución" = "ejecucion",
                    "En ejecución" = "ejecucion",
                    "En Ejecución" = "ejecucion",
                    "Enejecución"  = "ejecucion",
                    "43741"        = "ejecucion",
                    "finalizada" = "finalizada",
                    "Finalizada" = "finalizada",
                    "Planeada" = "planeada",
                    "planeada con financiamiento" = "planeada",
                    "planeada sin financiamiento" = "planeada"))) %>% 
  replace_na(list(estatus = "ejecucion")) %>% 
  # removing all planned activities 
  filter(estatus != "planeada")
```
:::

::: {.fold .s .o}
```{r location-cleaning}

locations <- read_excel("locations_20191111_1600.xlsx") %>%
  clean_names() %>% 
  rename(ubicacion = comunidad_o_centro) %>% 
  mutate(estado    = rm_accent(str_to_upper(estado)), # just to make sure 
         municipio = rm_accent(str_to_upper(municipio)),
         parroquia = rm_accent(str_to_upper(parroquia)),
         ubicacion = rm_accent(str_to_upper(ubicacion))) %>% 
  select(estado, pcode1, municipio, pcode2, parroquia, pcode3, ubicacion)

# next step is to resolve the admin locations
adm_dirty <- ven1 %>% 
  filter(is.na(estado) | is.na(pcode1) |
         is.na(municipio) | is.na(pcode2) |
         is.na(parroquia) | is.na(pcode3))

adm_clean <- ven1 %>% 
  filter(!is.na(estado) & !is.na(pcode1) &
         !is.na(municipio) & !is.na(pcode2) &
         !is.na(parroquia) & !is.na(pcode3)) 

# the distinct() actually clears out some duplicates in the data, 
#so it doesn't tally with the dataset from before 
# we'll get the rest of the duplicates in u_ben and act_ben
adm_dirty <- adm_dirty %>% 
  left_join(locations, by = "ubicacion") %>% 
  mutate(estado    = coalesce(estado.x, estado.y),
         pcode1    = coalesce(pcode1.x, pcode1.y),
         municipio = coalesce(municipio.x, municipio.y),
         pcode2    = coalesce(pcode2.x, pcode2.y),
         parroquia = coalesce(parroquia.x, parroquia.y),
         pcode3    = coalesce(pcode3.x, pcode3.y)) %>% 
  select(-estado.x, -estado.y, -pcode1.x, -pcode1.y,
         -municipio.x, -municipio.y, -pcode2.x, -pcode2.y,
         -parroquia.x, -parroquia.y, -pcode3.x, -pcode3.y) %>% 
  distinct() 
  
#overwriting ven1 to reduce environment objects
ven1 <-  bind_rows(adm_clean, adm_dirty)

```
:::

::: {.fold .s}
```{r outputs-cleaning}
# I'm kinda doubting the use of u_ben, ya I think take it out? since you're only using it once
# Am I just making these out of habit? I could cut them and make them inside the 
# code chunk for parr, but maybe I can find some justification for their existence, 
# maybe the disaggregations? 
u_ben <- ven1 %>% 
   # selecting out the very many disaggregation columns which we won't use
  select(-c(22:92)) %>% 
  pivot_longer(f_0_18:m_18plus, names_to = "desagregacion", values_to = "beneficiarios") %>% 
  filter(categoriadeactividad != "Vacunacion") %>% 
  filter(beneficiarios != 0) %>% 
  group_by(ubicacion, desagregacion) %>% 
  slice(which.max(beneficiarios)) %>% 
  ungroup()

act_ben <- ven1 %>% 
  select(-c(22:92)) %>% 
  pivot_longer(f_0_18:m_18plus, names_to = "desagregacion", values_to = "beneficiarios") %>% 
  filter(beneficiarios != 0) %>% 
  group_by(ubicacion, desagregacion, actividad) %>% 
  slice(which.max(beneficiarios)) %>% 
  ungroup()

rbind(sum(u_ben$beneficiarios), 
      sum(act_ben$beneficiarios), 
      sum(u_ben$beneficiarios) - sum(act_ben$beneficiarios))

```
:::

::: {.fold .s .o}
```{r making-all-parr}
# I think this is a gigantic chunk -- cannot decide if I would rather have less things in the 
# environment, or if I want more readable chunks. The benefit here is that if I want to change 
# something, I just have to change this chunk -- this also precludes the need to load cen_ref

parr <- u_ben %>% 
  group_by(pcode3, estado, municipio) %>% 
  summarise(beneficiarios = sum(beneficiarios)) %>% 
  ungroup() %>% 
  left_join(act_ben %>% 
             filter(categoria != "Vacunacion") %>%
             group_by(ubicacion, desagregacion, sector) %>% 
                slice(which.max(beneficiarios)) %>% 
                ungroup() %>% 
                pivot_wider(names_from = sector, values_from = beneficiarios) %>% 
                group_by(pcode3) %>% 
                # getting sector totals per pcode3
                summarise(educacion_ben = sum(Educacion, na.rm = TRUE),
                          nutricion_ben = sum(Nutricion, na.rm = TRUE),
                          proteccionGBV_ben = sum(Proteccion_GBV, na.rm = TRUE),
                          proteccionGeneral_ben = sum(Proteccion_General, na.rm = TRUE),
                          proteccionNNA_ben = sum(Proteccion_NNA, na.rm = TRUE),
                          salud_ben = sum(Salud, na.rm = TRUE),
                          seguridad_alimentaria_ben = sum(Seguridad_Alimentaria, na.rm = TRUE),
                          wash_ben = sum(WASH, na.rm = TRUE),
                          org_count = n_distinct(org_implementadora)) %>% 
             ungroup()) %>% 
  filter(str_detect(pcode3, "^VE")) %>% 
  # right_join to the census data
  right_join(read_excel("census data 20191122.xlsx", sheet = "data") %>% 
        clean_names() %>% 
        # selecting variables and renaming them with select
        select(estado, pcode1, municipio, pcode2, parroquia, pcode3, 
               fo = field_office,
               poblacion_2019 = x_2019_poblacion_parroquial_total,
               hogares_2011 = numero_de_hogares, 
               ham_2019_ambitos_ge, 
               percent_pobre = ham_2019_xx_pobreza_env_por_parroquia, 
               pob_pobre = ham_2019_xx_poblacion_pobre_por_parroquia, 
               poblacion_total_2011,
               poblacion_infantil_menor_de_12_anos, poblacion_adolescentes_de_12_a_17_anos,
               poblacion_de_18_anos_y_mas, 
               percent_urbana = poblacion_urbana_percent, 
               area_km2, 
               densidad_ppl_km2 = densidad_poblacional_ppl_km2,
               matricula_2017_educacion_inicial, matricula_2017_educacion_primaria, 
               matricula_2017_educacion_media, razon_de_dependencia_total,
               razon_de_dependencia_de_menores_de_15_anos, 
               percent_sin_agua_segura = x_abast_agua2_percent_sin_agua_segura,
               percent_sin_saneamiento_mejorado = x_saneamiento_percent_sin_saneamiento_mejorado,
               percent_analfabeto = percent_poblacion_10_anos_y_mas_analfabeta,
               promedio_de_personas_por_vivienda,
               percent_hogares_jefatura_femenina = percent_de_hogares_con_jefatura_femenina,
               percent_sin_servicio_electrico =
                 servicio_electrico_percent_no_tiene_servicio_electrico,
               ham_2019_x_violencia_envelope, ham_2019_x_mortalidad_y_salud_envelope, 
               ham_2019_x_pobreza_envelope, promedio_de_edad) %>% 
        mutate(estado    = rm_accent(str_to_upper(estado)), # just to make sure 
               municipio  = rm_accent(str_to_upper(municipio)),
               parroquia  = rm_accent(str_to_upper(parroquia))) %>% 
        # creating new disaggregation variables 
        mutate(pob_menor_de_18 = (poblacion_infantil_menor_de_12_anos +
                                 poblacion_adolescentes_de_12_a_17_anos) /poblacion_total_2011 *
                                 poblacion_2019, 
               pob_18_y_mas    = poblacion_de_18_anos_y_mas / poblacion_total_2011 * poblacion_2019, 
               hogares_2019    = hogares_2011 * poblacion_2019 / poblacion_total_2011, 
               matricula_total = matricula_2017_educacion_inicial + 
                                 matricula_2017_educacion_primaria + 
                                 matricula_2017_educacion_media) %>% 
        # dividing columns by 100 to clean then and put them between 0 and 1
        mutate_at(vars(percent_analfabeto, percent_sin_servicio_electrico, 
                       percent_sin_agua_segura,
                       percent_sin_saneamiento_mejorado,
                       percent_hogares_jefatura_femenina, percent_urbana,
                       razon_de_dependencia_total), ~(. / 100)) %>% 
        # mutating new columns with populations
        mutate(pob_analfabeto               = percent_analfabeto * poblacion_2019,
               pob_sin_agua_segura          = percent_sin_agua_segura * poblacion_2019, 
               pob_sin_servicio_electrico   = percent_sin_servicio_electrico * poblacion_2019,
               pob_sin_saneamiento_mejorado = percent_sin_saneamiento_mejorado * poblacion_2019,
               pob_urbana                   = percent_urbana * poblacion_2019) %>% 
        select(-c(matricula_2017_educacion_inicial, matricula_2017_educacion_primaria, 
               matricula_2017_educacion_media, poblacion_total_2011, hogares_2011,
               poblacion_infantil_menor_de_12_anos, poblacion_adolescentes_de_12_a_17_anos, 
               poblacion_de_18_anos_y_mas)),
             by = "pcode3") %>% 
  # mutating new variables and making sure NAs become 0s 
  mutate(beneficiarios          = ifelse(is.na(beneficiarios), 0, beneficiarios),
         org_count              = ifelse(is.na(org_count), 0, org_count),
         not_covered_pobre      = pob_pobre - beneficiarios,
         coverage_percent       = beneficiarios / poblacion_2019,
         coverage_pobre_percent = beneficiarios / pob_pobre,
         percent_total_ben      = beneficiarios / sum(beneficiarios),
         org_present            = ifelse(beneficiarios > 0, TRUE, FALSE),
         pob_pobre_score     = rescale(pob_pobre, to = c(0,1)), 
         percent_pobre_score = rescale(percent_pobre, to = c(0,1)), 
         poverty_score       = (pob_pobre_score + percent_pobre_score) / 2) %>% 
  select(-c(estado.x, municipio.x)) %>% 
  rename(estado = estado.y, 
         municipio = municipio.y)

```
:::

::: {.fold .s}
```{r}
parr %>% 
  mutate(coverage_type = case_when(not_covered_pobre <= 0 ~ "over",
                                   not_covered_pobre > 0 & beneficiarios >= 1 ~ "under", 
                                   beneficiarios == 0 ~ "not_covered")) %>% 
  group_by(coverage_type) %>% 
  summarise(parroquias = n(),
            beneficiarios = sum(beneficiarios),
            avg_org_count = mean(org_count),
            percent_pobre = (sum(pob_pobre)) / (sum(poblacion_2019)),
            percent_urbana = (sum(pob_urbana)) / (sum(poblacion_2019)),
            percent_sin_agua_segura = (sum(pob_sin_agua_segura)) / (sum(poblacion_2019)),
            percent_sin_saneamiento_mejorado = (sum(pob_sin_saneamiento_mejorado)) /
              (sum(poblacion_2019))) %>% 
  gather(key = var_name, value = value, 2:ncol(.)) %>% 
  spread_(key = names(.)[1], value = 'value') %>% 
  relocate(not_covered, .after = under) %>% 
  pander(big.mark = ",")

parr %>% 
  mutate(coverage_type = case_when(not_covered_pobre <= 0 ~ "over",
                                   not_covered_pobre > 0 & beneficiarios >= 1 ~ "under", 
                                   beneficiarios == 0 ~ "not_covered")) %>% 
  filter(coverage_type == "over") %>% 
  select(estado, municipio, estado, parroquia, beneficiarios, pob_pobre) %>%
  mutate(coverage_percent = beneficiarios / pob_pobre * 100) %>% 
  arrange(desc(beneficiarios)) %>% 
  pander(big.mark = ",")
  
```
:::

> There are some clear data entry problems with the parroquia of Altagracia in Distrito Capital -- likely, partners have reported activities that occurred in other parts of the capital all in one parroquia as the total number of benefificaries reached in the whole of Distrito Capital is `r summarise(parr, beneficiarios = sum(beneficiarios[municipio == "LIBERTADOR"]))`. 

```{r}
parr %>% 
  filter(beneficiarios != 0) %>% 
  ggplot(aes(x = densidad_ppl_km2, y = beneficiarios)) +
  geom_point() +
  geom_smooth(method = "lm") +
  scale_y_continuous(trans = "log10", labels = comma) +
  scale_x_continuous(trans = "log10") +
  stat_regline_equation(label.y = 1e-01, aes(label = ..rr.label..))

# nothing I've tried seems to have a high r-squared 
lm(beneficiarios ~ percent_urbana, data = 
                 parr %>% 
                 filter(beneficiarios <= 5000)) %>% 
  summary()

```

> Several trees have been built to split parroquias up into targetting groups based on their 
characteristics. The variables selected largely originate from the census dataset, with some having been upgraded by the 2019 Municipal Prioritisation Tool, which was a Principal Components Analysis of key variables related to poverty, health and mortality and violence and insecurity. 

> The specific variables and formulae used for each of the trees can be seen by unhiding the source code below. After numerous tests 

```{r trees}
# the trees are all overfitted, but that's ok, we're just trying to organise where to go next 
# based on the data we have at the moment. We will fit a new tree as more data comes in -- 
# it doesn't make that much sense to try and predict where agencies will expand.

set.seed(3000)

parr0 <- parr %>% filter(not_covered_pobre >= 1)

# number of not covered poor persons 
tree1 <- parr0 %>%
  rpart(not_covered_pobre ~ estado + percent_pobre + percent_urbana + 
        densidad_ppl_km2 + razon_de_dependencia_de_menores_de_15_anos + razon_de_dependencia_total +  
        percent_sin_agua_segura + percent_sin_saneamiento_mejorado +
        percent_sin_servicio_electrico + percent_analfabeto + percent_hogares_jefatura_femenina, 
        promedio_de_personas_por_vivienda, data = ., minbucket = 100)

# interested again -- just to show the decision tree of how 
tree2 <- parr0 %>% 
  rpart(beneficiarios ~ percent_pobre + percent_urbana + densidad_ppl_km2 + 
        razon_de_dependencia_de_menores_de_15_anos + razon_de_dependencia_total +  
        percent_sin_agua_segura + percent_sin_saneamiento_mejorado +
        percent_sin_servicio_electrico + percent_analfabeto + percent_hogares_jefatura_femenina, 
        promedio_de_personas_por_vivienda, data = ., 
        control = rpart.control(minsplit = 2, minbucket = 100))

        
# tree based on poverty_score
tree3 <- parr0 %>%
  rpart(poverty_score ~ estado + percent_urbana + densidad_ppl_km2 + 
        razon_de_dependencia_de_menores_de_15_anos + razon_de_dependencia_total +  
        percent_sin_agua_segura + percent_sin_saneamiento_mejorado +
        percent_sin_servicio_electrico + percent_analfabeto + percent_hogares_jefatura_femenina, 
        promedio_de_personas_por_vivienda, data = ., minbucket = 150)

parr0 <- parr0 %>% 
  mutate(rule1 = row.names(tree1$frame)[tree1$where]) %>%
      left_join(rpart.rules.table(tree1) %>% 
      filter(Leaf == TRUE) %>% 
      rename(rule1 = Rule) %>% 
      group_by(rule1) %>% 
      summarise(subrules1 = paste(Subrule, collapse = ",")))  %>% 
  mutate(rule3 = row.names(tree3$frame)[tree3$where]) %>%
      left_join(rpart.rules.table(tree3) %>% 
      filter(Leaf == TRUE) %>% 
      rename(rule3 = Rule) %>% 
      group_by(rule3) %>% 
      summarise(subrules3 = paste(Subrule, collapse = ","))) %>% 
  mutate(rule3a = recode(as.character(rule3),
                         `4` = "A", `11` = "B", `10` = "C", `6` = "D", `7` = "E"),
         rule1a = recode(as.character(rule1),
                         `7` = "V", `6` = "W", `11` = "X", `10` = "Y", `4` = "Z"))
```

```{r}
fancyRpartPlot(tree1, digits = -3, sub = "", palettes = "Blues", type = 2)
fancyRpartPlot(tree2, digits = -3, sub = "", palettes = "Blues", type = 2)
fancyRpartPlot(tree3, digits = -3, sub = "", palettes = "Blues", type = 2)
```


```{r}
# for showing the output of the tree nicely
as.party(tree3)

printcp(tree2)

```

> Tree1 splits parroquias by

::: {.fold .s .o}
```{r tree1-notes}
# figure out the right order for the leaves
parr0 %>% 
  group_by(rule1a) %>% 
  summarise(parr_no_ben = n_distinct(pcode3[beneficiarios == 0]), 
            beneficiarios = sum(beneficiarios),
            ben_per_parr = sum(beneficiarios) / n(), 
            not_covered = sum(not_covered_pobre),
            nc_per_parr = sum(not_covered_pobre) / n(),
            nc_per_mun = sum(not_covered_pobre) / n_distinct(pcode2), 
            avg_org_count = mean(org_count),
            coverage_percent = sum(beneficiarios) / sum(poblacion_2019),
            percent_pobre = sum(pob_pobre) / sum(poblacion_2019),
            percent_urbana = sum(pob_urbana) / sum(poblacion_2019),
            densidad_ppl_km2 = sum(poblacion_2019) / sum(area_km2, na.rm = TRUE),
            parroquias = n(),
            municipios = n_distinct(pcode2),
            parr_per_mun = n() / n_distinct(pcode2)) %>% 
  gather(key = var_name, value = value, 2:ncol(.)) %>% 
  spread_(key = names(.)[1], value = 'value') %>% 
  arrange(factor(var_name, levels = c("not_covered", "nc_per_parr", "nc_per_mun", "beneficiarios",
                                      "ben_per_parr",  "avg_org_count", "coverage_percent",
                                      "percent_pobre", "percent_urbana", "densidad_ppl_km2", 
                                      "parroquias", "parr_no_ben", "municipios", 
                                      "parr_per_mun"))) %>%  pander(big.mark = ",")


# V is dense, urban and highest operational presence, shortly followed by W
# Z is a priority for reaching the most vulnerable and marginalised, 
# but it truly is very sparsely populated. The number of persons you can reach is low and 
# this is the most operationally challenging
# Y is more than 50% poor, but also sparsely populated; but it has more not_covered than X,  with double the number of not_covered per pcode3, it also has less than half the 
# municipalities -- operationally more feasible to move into this area
# X is spread out, numerous and you should expand there only if you have operations in adjacent areas

```
:::

```{r tree3-rules}
# figure out the right order for the leaves
# I'm liking tree3 more and more
  
parr0 %>% 
  group_by(rule3a) %>% 
  summarise(parr_no_ben = n_distinct(pcode3[beneficiarios == 0]),
            beneficiarios = sum(beneficiarios),
            ben_per_parr = sum(beneficiarios) / n(), 
            not_covered = sum(not_covered_pobre),
            nc_per_parr = sum(not_covered_pobre) / n(),
            nc_per_mun = sum(not_covered_pobre) / n_distinct(pcode2),
            avg_org_count = mean(org_count),
            coverage_percent = sum(beneficiarios) / sum(poblacion_2019),
            percent_pobre = sum(pob_pobre) / sum(poblacion_2019),
            percent_urbana = sum(pob_urbana) / sum(poblacion_2019),
            densidad_ppl_km2 = sum(poblacion_2019) / sum(area_km2, na.rm = TRUE),
            parroquias = n(),
            municipios = n_distinct(pcode2),
            parr_per_mun = n() / n_distinct(pcode2)) %>% 
  gather(key = var_name, value = value, 2:ncol(.)) %>% 
  spread_(key = names(.)[1], value = 'value') %>% 
  arrange(factor(var_name, levels = c("not_covered", "nc_per_parr", "nc_per_mun", "beneficiarios",
                                      "ben_per_parr",  "avg_org_count", "coverage_percent",
                                      "percent_pobre", "percent_urbana", "densidad_ppl_km2", 
                                      "parroquias", "parr_no_ben", "municipios", 
                                      "parr_per_mun"))) %>%  pander(big.mark = ",")


# A is population centres which are easy to reach, but with only 24% of the population being poor, 
# careful targetting and beneficiary selection is required. 
# B is probably the best option for expansion -- it has the highest concentration of unconvered 
# persons per parroquia and municipio, is substantially poorer than A (at 41% poverty incidence). 
# Operational expansion is more likely, due to the higher concentration of organisations and 
# it is also fairly urban, meaning that this uncovereved population is fairly easy to reach. 
# better than C in almost every way -- higher concentration of NC, less area to cover and 
# it's already got higher operational coverage, so that makes; 
# in fact, B should be prioritised here -- it's easy to 
# in many ways, D is better than C as well
# E is just a giant operational challenge

```

```{r}
# let's try and do principal component analysis on u_parr --
# one PC for WASH, one for urban/pop_den
# and let's remove promedio por edad
```

```{r state-plot}
parr0 %>% 
  group_by(estado) %>% 
  summarise(not_covered_pobre = sum(not_covered_pobre),
            poverty_rate = sum(pob_pobre) / sum(poblacion_2019) * 100) %>% 
  ggplot(aes(x = fct_reorder(estado, not_covered_pobre, .desc = TRUE), 
             y = not_covered_pobre, 
             fill = poverty_rate))+
  scale_fill_continuous(type = "gradient", trans = "reverse") +
  geom_col() + 
  geom_text(aes(label = scales::comma(not_covered_pobre, accuracy = 1)), 
            vjust = -0.25, hjust = 0.05, size = 2, angle = 60) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.key.size = unit(0.5, "cm"),
        legend.text = element_text(size = 7),
        legend.title = element_text(size = 7)) +
  xlab("") + ylab("Poor persons not covered") +
  scale_y_continuous(labels = comma, limits = c(0, 2000000))
  
```

```{r}
pcode3_shape <- st_read("C:/Users/Sean Ng/Documents/R/coverage_gaps_venezuela/ven_admbnda_adm3_20180502/ven_admbnda_adm3_20180502.shp",
                        quiet = TRUE) %>% 
  rename(pcode1 = ADM1_PCODE,
         pcode2 = ADM2_PCODE,
         pcode3 = ADM3_PCODE)

```
